Precision calculations for cosmic shear

### 2017-02-08 discussion.tex
Updated the alternative two-point discussion section

\subsection{Comparison of two-point shear statistics; the two-point correlation
function, mass aperture statistic and COSEBIs}

\ch{Copied this over from original comment paper - I would like to include this
material somewhere in the paper - this needs work though to make it fit....
TBD}

\citet{CFHTLenS-2pt-notomo} present a detailed comparison of cosmological
constraints obtained from a range of different two-point shear statistics
including the shear correlation function, $\xi_\pm$, the aperture-mass
dispersion, $\langle M_ {\rm ap}^2 \rangle2$ \citep{1998MNRAS.296..873S}, and
the COSEBIs, $E_n$ \citep{COSEBIs}.  These statistics are linearly related to
the shear power spectrum in the flat-sky approximation via integrals of the form (omitting tomographic indices),
%
\begin{align}
  \langle M_ {\rm ap}^2 \rangle(\theta) & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi}
  \hat U^2(\theta\ell) P^\gamma(\ell),
  \label{eqn:integ}
  \\
  E_n & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P^\gamma(\ell),
\end{align}
%
where the Fourier-space filter functions $\hat U$ and $W_n$ are defined in
\cite{1998MNRAS.296..873S} and \cite{COSEBIs}. The corresponding equations for
$\xi_\pm$ are given in equation~\ref{eqn:xiGG}. Figure~\ref{fig:filters} shows
the integrands of these statistics for two cases, normalised to their maximum
value, where the integrands are of the form $\ell F(\ell) P^\gamma(\ell)$. For
$\xi_\pm$, $P^\gamma(\ell)$ is equal to the sum of the E and B-mode power spectra
which cannot easily be separated,
motivating the development of the aperture-mass dispersion statistic, $\langle
M_ {\rm ap}^2 \rangle$. This statistic is, however, a lossy conversion and is
biased by small angular separations, where blending of galaxies makes shear
measurement challenging \citep{KSE06}. The COSEBIs statistic tackles both
these shortcomings.  The upper two panels of figure~\ref{fig:filters} show the
integrands\footnote{Note that the lower left panel of Fig.~1 in
\citet{2016arXiv161104954K} shows an integrated form of this function for a
maximum angular separation of $100^\prime$. However, in
\citet{CFHTLenS-2pt-notomo}, the data used in \citet{2016arXiv161104954K},
$\theta$ is between $0.8'$ and $350'$.} of $\xi_\pm$ for $\theta=100'$ and
$\theta=350'$. The lower middle panel in Figure~\ref{fig:filters} shows the
COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where
we only show the integrands for the lowest COSEBIs mode, $E_1$, as the higher
modes generally probe larger $\ell$-modes.  Finally, the lowest panel shows the
integrands of aperture mass dispersion statistics, for the same two maximum
angular ranges.

\begin{figure}%[!htp]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.48\textwidth]{figures/IntegAll.pdf} \\
\end{tabular}
\caption{ \small{\label{fig:filters}. Integrand of $\xi_+$ (upper), $\xi_-$
(upper middle), $E_1$ (lower middle, E-COSEBIs) and $\langle M_{\rm ap}
\rangle^2$ (lower panel). All integrands are of the form $\ell F(\ell)
P(\ell)$, where $F(\ell)$ is the corresponding weight-function for each
statistic and $P(\ell)$ is the E-mode convergence power spectrum, with the
exception of $\xi_\pm$, for which $P(\ell)$ is equal to the sum of the E and
B-mode power spectra. Two cases are shown for each statistic as listed in each
caption. For the aperture mass statistic $\theta_{\rm max}=2\theta$ is shown.
Note that higher order COSEBIs modes generally probe larger $\ell$-modes, hence
here we only show the lowest mode $E_1$. All values are normalized with respect
to their maximum value. } }
\end{center}
\end{figure}

From Figure~\ref{fig:filters} we can see that the two-point cosmic shear
statistics tested by \citet{CFHTLenS-2pt-notomo} exhibit different dependences
between the angular scales sampled and the $\ell$-range probed.   If a
significant bias had been introduced at low-$\ell$ by using flat-sky and Limber
approximations, we would then expect to see a systematic shift between the
different two-point statistics with the COSEBIs statistic being essentially
unaffected as it only includes modes with $\ell \gtrsim 20$.  This is found not
to be the case with all three statistics finding $\sigma_8
(\Omega_m/0.27)^\alpha = 0.79$ with errors that range from $0.03$ to $0.06$ for
the full five-parameter fit, and $\alpha$ ranging from $0.59$ to 0.7 \citep[see
Table 5 of][]{CFHTLenS-2pt-notomo}.  This comparison further supports our
argument that the approximations highlighted by \citet{2016arXiv161104954K}
have negligible impact for current surveys.

>>>>

\subsection{Alternative two-point shear statistics; the mass aperture statistic and COSEBIs}

The two-point shear correlation function $\xi_\pm$ represents the current baseline observable for cosmic shear measurements.   
As shown in Figure~\ref{fig:xi_pm_K16}, however, using the standard first-order extended Limber flat-sky approximation (ExtL1FlHyb) can result
in errors exceeding 10 percent, on angular scales $\theta > 300$ arcmin.    This is a result of the significant weight given to low $\ell$ modes in the $\xi_+$ statistic, as illustrated in Figure~\ref{fig:filters} which shows the integrand of $\xi_+$ and $\xi_-$ (upper two panels) for two cases ( $\theta = 100$ and $\theta = 350$ arcmin), normalised to their maximum
value.   This error does not impact CFHTLenS analyses, given the low signal-to-noise of the measurements on these scales.  It will however become increasingly important for upcoming wider-field surveys which will accurately probe these scales.

In this paper we provide a solution in the form of the second-order extended Limber approximation, but another option is to consider alternative two-point shear statistics that are less sensitive to accuracy in shear power spectrum measurement at low $\ell$.   Both the aperture-mass dispersion, $\langle M_ {\rm ap}^2 \rangle2$ \citep{1998MNRAS.296..873S}, and
the Complete Orthogonal Sets of E/B-mode Integrals (COSEBIs), $E_n$ \citep{COSEBIs} statistics satisfy this requirement and are linearly related to
the shear power spectrum in the flat-sky approximation via integrals of the form (omitting tomographic indices),
%
\begin{align}
  \langle M_ {\rm ap}^2 \rangle(\theta) & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi}
  \hat U^2(\theta\ell) P^\gamma(\ell),
  \label{eqn:integ}
  \\
  E_n & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P^\gamma(\ell),
\end{align}
%
where the Fourier-space filter functions $\hat U$ and $W_n$ are defined in
\cite{1998MNRAS.296..873S} and \cite{COSEBIs}. 
Figure~\ref{fig:filters} shows the integrands of these statistics, again normalised to their maximum
value, where the integrands are of the form $\ell F(\ell) P^\gamma(\ell)$.  
The lower middle panel in Figure~\ref{fig:filters} shows the
COSEBIs integrands for two angular ranges, $[1',100']$ and $[0.8',350']$, where
we only show the integrands for the lowest COSEBIs mode, $E_1$, as the higher
modes generally probe larger $\ell$-modes.  The lowest panel shows the
integrands of aperture mass dispersion statistics, for the same two maximum
angular ranges.

Note that the development of the aperture-mass dispersion statistic, $\langle
M_ {\rm ap}^2 \rangle$ was initially motivated to enable the separation of the measured signal into an 
E-mode (cosmological signal) and B-mode (systematics). This statistic is, however, a lossy conversion and is
biased by small angular separations, where blending of galaxies makes shear
measurement challenging \citep{KSE06}. The COSEBIs statistic tackles both
these shortcomings.  \citet{CFHTLenS-2pt-notomo} present a detailed comparison of cosmological
constraints obtained from this range of different two-point shear statistics finding consistent results.

\begin{figure}%[!htp]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.48\textwidth]{figures/IntegAll.pdf} \\
\end{tabular}
\caption{ \small{\label{fig:filters}. 
Integrand of $\xi_+$ (upper), $\xi_-$
(upper middle), $E_1$ (lower middle, E-COSEBIs) and $\langle M_{\rm ap}
\rangle^2$ (lower panel). All integrands are of the form $\ell F(\ell)
P(\ell)$, where $F(\ell)$ is the corresponding weight-function for each
statistic and $P(\ell)$ is the E-mode convergence power spectrum, with the
exception of $\xi_\pm$, for which $P(\ell)$ is equal to the sum of the E and
B-mode power spectra. Two cases are shown for each statistic as listed in each
caption. For the aperture mass statistic $\theta_{\rm max}=2\theta$ is shown.
Note that higher order COSEBIs modes generally probe larger $\ell$-modes, hence
here we only show the lowest mode $E_1$. All values are normalized with respect
to their maximum value. This figure illustrates how different two-point cosmic shear
statistics have different dependences between the angular scales sampled and the $\ell$-range probed. }}
\end{center}
\end{figure}
